---
title: 'Assignment: Course Project 1'
author: "Thiago Augusto Cardoso"
date: "June 8, 2016"
output: html_document
---

For producing a fully reproducible document, first we load the required libraries and set the working directory

```{r libraries, echo=TRUE, include=TRUE,message=FALSE,cache=TRUE}
## load required libraries
library(dplyr)
library(RCurl)
library(rio)
library(lubridate)
library(ggplot2)

## set working directory
setwd('/home/mascc2/coursera/reproducible-research')

```

In a second step, we actually download the data from source at [cloud front, acessed in 2016-06-08] ('https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip') and store in a data.frame

```{r download , cache=TRUE}
## download data and store in a variable
activity<-data.frame()
activity<-import('https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip')
``` 
For assuring that the data actually loaded properly we echo the variable

```{r echo, cache=TRUE}
head(activity)
```

##1. We also load the data with the dplyr package for easier filtering and aggregation.

```{r dplyr , cache=TRUE}
## load file 
activitydplyr<-tbl_df(activity)
```

## Then we procede with the aggregation to have the number of steps by date.

```{r aggregation , cache=TRUE}
## generate the number of steps per day
stepsbydate<-aggregate(steps ~ date, activitydplyr,sum,na.rm=TRUE)
```

## 2. Now we generate the histogram of steps per day according to question number 2

```{r histogram , cache=TRUE}
## generate histogram of steps per day
hist(stepsbydate$steps,main='Histogram of steps per day',xlab='Number of Steps')

```

## 3. Now we generate the mean and median number of steps taken each day according to question number 3

Calculating the mean steps per day
```{r mean, cache=TRUE,echo=TRUE}
## take the mean number of steps taken each day
meanstepsbydate <-aggregate(steps ~ date, activitydplyr,mean,na.rm=TRUE)
```

```{r echomean, cache=TRUE,echo=TRUE}
meanstepsbydate
```

Calculating the median steps per day
```{r median, cache=TRUE, echo=TRUE}
medianstepsbydate <-aggregate(steps ~ date, activitydplyr,median,na.rm=TRUE)
```

```{r echomedian, cache=TRUE,echo=TRUE}
medianstepsbydate
```

## 4. Times series plot of the average number of steps taken each day

Generating the time series plot

```{r plotmean, cache=TRUE,echo=TRUE}
meanstepsbydate$date<-as.Date(meanstepsbydate$date)
plot(meanstepsbydate$date,meanstepsbydate$steps,xlab='Date',ylab = 'Mean number of steps per day',main="Average number of steps taken",type='l') 
```

## 5. The 5-minute interval that, on average, contains the maximum number of steps

Selecting the maximum number of steps on average

```{r maxsteps, cache=TRUE,echo=TRUE}
meanstepsbydatedplyr<-tbl_df(meanstepsbydate)
max<-head(arrange(meanstepsbydatedplyr,desc(steps)),1)
max
```

## 6. Code to describe and show a strategy for imputing missing data

Show the number of missing cases

```{r missingcases, cache=TRUE, echo=TRUE}
missingcases <- sum(is.na(activitydplyr$steps))
missingcases
```

Inputting missing cases with the mean

```{r nonmissing, cache=TRUE,echo=TRUE}
activitynonna <- activitydplyr
activitynonna$steps[is.na(activitynonna$steps)] <- mean(activitynonna$steps, na.rm=TRUE)
activitynonnaperday <- aggregate(steps ~ date, data=activitynonna, sum, na.rm=TRUE)
activitynonna
str(activitynonna)
```

## 7. Histogram of the total number of steps taken each day after missing values are imputed
Generating the histogram 
```{r histnonna, cache=TRUE,echo=TRUE}
hist(activitynonnaperday$steps,main='Histogram of steps per day',xlab='Number of Steps')
```

## 8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

Create a function to check if there is a weekday
```{r function,cache=TRUE,echo=TRUE}
is_weekday <-function(date){
        if(wday(date) %in% c(1,7)) result<-"weekend"
        else
                result<-"weekday"
        result
}
```

create the variable weekday and check how many weekdays
```{r activitynonna, cache=TRUE,echo=TRUE}
activitynonna <- mutate(activitynonna,date=ymd(date)) %>% mutate(weekday=sapply(date,is_weekday))

table(activitynonna$weekday)
```

plot the graph
```{r plotpanel,cache=TRUE}
daily_patterns <- activitynonna %>% mutate(weekday=factor(weekday,levels=c("weekend","weekday"))) %>% 
group_by(interval,weekday)

plot(daily_patterns$weekday,daily_patterns$steps)
```

